On the validity of the Born approximation for beyond-Gaussian weak lensing observables

### 2017-05-17 Paper/paper.tex
round 4 of referee comments

\begin{figure*}
\begin{center}
\includegraphics[scale=0.45]{Figures/bias_ngal_convergence_powerSN_s0_nb100.eps} \includegraphics[scale=0.45]{Figures/bias_ngal_convergence_momentsSN_s50_nb9.eps}
\end{center}
\caption{Parameter bias, obtained from the $\kappa$ power spectrum \textbf{\color{red} (left panel) and the $\kappa$ moments (right panel)}, as a function of the galaxy density angular $n_g$ used to compute the shape noise rms in eq. (\ref{par:shaperms}). The plotted quantity is the difference between the mean of 1000 parameter fits using the estimator (\ref{par:fisherest}) with $\bb{M}$ measured with and without the Born approximation. The difference is shown in units of the estimator standard deviation. We show the trend for $\Omega_m$ (blue), $w_0$ (red) and $\sigma_8$ (green).}
\label{fig:parbiasSN}
\end{figure*}

@@

Figure \ref{fig:psResiduals} shows a comparison between the Born--approximated $\kappa$ power spectra and the ones obtained with full ray--tracing, separating different orders in $\Phi$. We can conclude that the dominant contributions to the residuals come from the Born--geodesic cross power, which dominates the Born--lens--lens cross power by two orders of magnitude on small scales. Figure \ref{fig:skResiduals} shows that the geodesic and lens--lens contributions to the $\kappa$ skewness and kurtosis are comparable and that the lowest non--trivial post--Born corrections can fully account for the discrepancy between the Born--approximated quantities and the fully ray--traced ones. The geodesic and lens--lens corrections to the skewness we find agree, as orders of magnitude, with the ones already published in \citep{WLBispectrumDodelson}, although our measured reduced shear correction is larger. \textbf{\color{red} We stress, however, that our estimates and the ones in \citep{WLBispectrumDodelson} have been produced under different assumptions. In particular, \citep{WLBispectrumDodelson} use a fully analytical approach based on the Limber approximation, they consider different smoothing window functions and use their own assumptions for the bi--spectrum of $\kappa$.} Figs. \ref{fig:psResiduals} and \ref{fig:skResiduals} indicate that truncating $\kappa$ at $O(\Phi^2)$ post--Born terms, leads to inaccurate predictions of the power spectrum and moments. In particular, higher-order terms involving $\langle \kappa_{\rm born}^n \kappa_{\rm ll}^m \kappa_{\rm geo}^2\rangle$ and higher, cause large deviations (as seen by the difference between the solid and dashed green curves) that, in an exact approach, are canceled by $O(\Phi^3)$ corrections to $\kappa$, as pointed out by \citep{HirataKrause,CMBPrattenLewis,ScoccimarroLoop}. \textbf{\color{blue} For example, regarding the skewness of $\kappa$, the residuals $\kappa^3_{\rm born+geo}-\kappa^3_{\rm born}$ contain terms up to order $O(\Phi^6)$. Some of these high order terms should be canceled by $O(\Phi^3)$ corrections to the $\kappa$ image. These cancellations do not show up if one truncates the expansion of $\kappa$ to $O(\Phi^2)$. This is the reason behind the discrepancy between the green curves in Figure \ref{fig:skResiduals}.} Therefore, in practice, in order to go beyond Born approximation, it is more advantageous, both for accuracy and for CPU time, to perform full ray--tracing (whose computational cost is comparable to those of the $O(\Phi^2)$ line--of--sight integrals; Table \ref{tab:benchmarks}).

@@

\section*{Acknowledgments}
We thank Uro\v{s} Seljak and Colin Hill for useful discussions. The $N$--body simulations in this work were performed at National Energy Research Scientific Computing Center (NERSC). We thank the LSST Dark Energy Science Collaboration (DESC) for the allocation of time. The ray--tracing simulations and parameter estimation calculations were performed at the NSF XSEDE facility, supported by grant number ACI-1053575, and at the New York Center for Computational Sciences, a cooperative effort between Brookhaven National Laboratory and Stony Brook University, supported in part by the State of New York. This work was supported in part by the U.S. Department of Energy under Contract No. DE-SC00112704, the NSF Grant No. AST-1210877 (to Z.H.), the Research Opportunities and Approaches to Data Science (ROADS) program at the Institute for Data Sciences and Engineering at Columbia University (to Z.H.) and a Simons Fellowship in Theoretical Physics (to Z.H.).

>>>>

\begin{figure*}
\begin{center}
\includegraphics[scale=0.45]{Figures/bias_ngal_convergence_powerSN_s0_nb100.eps} \includegraphics[scale=0.45]{Figures/bias_ngal_convergence_momentsSN_s50_nb9.eps}
\end{center}
\caption{Parameter bias, obtained from the $\kappa$ power spectrum \textbf{\color{red} (left panel) and the $\kappa$ moments (right panel)}, as a function of the galaxy density angular $n_g$ used to compute the shape noise rms in eq. (\ref{par:shaperms}). The plotted quantity is the difference between the mean of 1000 parameter fits using the estimator (\ref{par:fisherest}) with $\bb{M}$ measured with and without the Born approximation. The difference is shown in units of the estimator standard deviation. We show the trend for $\Omega_m$ (blue), $w_0$ (red) and $\sigma_8$ (green). \textbf{\color{ForestGreen} The quoted $\sigma_{\rm ray}$ is calculated as a standard deviation. Parameter estimates PDFs are not Gaussian because the estimator $\hat{\mathbf{d}}_{\rm obs}$ which appears in equation (\ref{par:fisherest}) in general is non--Gaussian distributed. When we average over multiple fields of view though (this is the case for LSST), the Gaussianity assumption is justified by the central limit theorem. This is why the standard deviation $\sigma_{\rm ray}$ is a reasonable benchmark for statistical significance.}}
\label{fig:parbiasSN}
\end{figure*}

@@

Figure \ref{fig:psResiduals} shows a comparison between the Born--approximated $\kappa$ power spectra and the ones obtained with full ray--tracing, separating different orders in $\Phi$. We can conclude that the dominant contributions to the residuals come from the Born--geodesic cross power, which dominates the Born--lens--lens cross power by two orders of magnitude on small scales. \textbf{\color{ForestGreen} The dashed blue line in Figure \ref{fig:psResiduals} corresponds to the residuals $P_{\rm ray} - P_{\rm born} - 2P_{\rm born,ll} - 2P_{\rm born,geo}$, which are comparable to the $O(\Phi^3)$ terms themselves. This means that the $O(\Phi^4)$ post--Born terms are completely overshot by cosmic variance and numerical noise in this case.} Figure \ref{fig:skResiduals} shows that the geodesic and lens--lens contributions to the $\kappa$ skewness and kurtosis are comparable and that the lowest non--trivial post--Born corrections can fully account for the discrepancy between the Born--approximated quantities and the fully ray--traced ones. The geodesic and lens--lens corrections to the skewness we find agree, as orders of magnitude, with the ones already published in \citep{WLBispectrumDodelson}, although our measured reduced shear correction is larger. \textbf{\color{red} We stress, however, that our estimates and the ones in \citep{WLBispectrumDodelson} have been produced under different assumptions. In particular, \citep{WLBispectrumDodelson} use a fully analytical approach based on the Limber approximation, they consider different smoothing window functions and use their own assumptions for the bi--spectrum of $\kappa$.} Figs. \ref{fig:psResiduals} and \ref{fig:skResiduals} indicate that truncating $\kappa$ at $O(\Phi^2)$ post--Born terms, leads to inaccurate predictions of the power spectrum and moments. In particular, higher-order terms involving $\langle \kappa_{\rm born}^n \kappa_{\rm ll}^m \kappa_{\rm geo}^2\rangle$ and higher, cause large deviations (as seen by the difference between the solid and dashed green curves) that, in an exact approach, are canceled by $O(\Phi^3)$ corrections to $\kappa$, as pointed out by \citep{HirataKrause,CMBPrattenLewis,ScoccimarroLoop}. \textbf{\color{blue} For example, regarding the skewness of $\kappa$, the residuals $\kappa^3_{\rm born+geo}-\kappa^3_{\rm born}$ contain terms up to order $O(\Phi^6)$. Some of these high order terms should be canceled by $O(\Phi^3)$ corrections to the $\kappa$ image. These cancellations do not show up if one truncates the expansion of $\kappa$ to $O(\Phi^2)$. This is the reason behind the discrepancy between the green curves in Figure \ref{fig:skResiduals}.} Therefore, in practice, in order to go beyond Born approximation, it is more advantageous, both for accuracy and for CPU time, to perform full ray--tracing (whose computational cost is comparable to those of the $O(\Phi^2)$ line--of--sight integrals; Table \ref{tab:benchmarks}).

@@

\section*{Acknowledgments}
\textbf{\color{ForestGreen} We thank the anonymous referee for the useful and insightful comments}. We thank Uro\v{s} Seljak and Colin Hill for useful discussions. The $N$--body simulations in this work were performed at National Energy Research Scientific Computing Center (NERSC). We thank the LSST Dark Energy Science Collaboration (DESC) for the allocation of time. The ray--tracing simulations and parameter estimation calculations were performed at the NSF XSEDE facility, supported by grant number ACI-1053575, and at the New York Center for Computational Sciences, a cooperative effort between Brookhaven National Laboratory and Stony Brook University, supported in part by the State of New York. This work was supported in part by the U.S. Department of Energy under Contract No. DE-SC00112704, the NSF Grant No. AST-1210877 (to Z.H.), the Research Opportunities and Approaches to Data Science (ROADS) program at the Institute for Data Sciences and Engineering at Columbia University (to Z.H.) and a Simons Fellowship in Theoretical Physics (to Z.H.).
